---
title: "Assignment2"
author: "Joshua Registe"
date: "9/4/2020"
output: html_document
---

Problem Set 1

(1) Show that AT A 6= AAT in general. (Proof and demonstration.)

The following example will compute a Matrix A, and the transpose of that matrix and check for equality
```{r}
library(matlib)

A<-matrix(c(1,7,4,63,6,7,2,9,67),nrow = 3)
A_T<-t(A)

A_T%*%A == A%*%A_T

```
The example shows that AT*A is not equal to A*AT

(2) For a special type of square matrix A, we get AT A = AAT. Under what conditions
could this be true? (Hint: The Identity matrix I is an example of such a matrix).

The condition will be true if the matrices are symstrical or identity matrices.

```{r}

A<- matrix(c(3,1,2,1,3,1,2,1,3),nrow =3)
A

A_T<-t(A)


A_T%*%A == A%*%A_T
```



Problem Set 2

Matrix factorization is a very important problem. There are supercomputers built just
to do matrix factorization. Every second you are on an airplane, matrices are being
factorized. Radars that track flights use a technique called Kalman filtering. At the heart
of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear
systems of equations when they track your flight using radars.
Write an R function to factorize a square matrix A into LU or LDU, whichever you
prefer. 

You don’t have to worry about permuting rows of A and you can assume that A is less
than 5x5, if you need to hard-code any variables in your code. If you doing the entire
assignment in R, then please submit only one markdown document for both the problems.


We create function and return an error if the matrix is too large (greater than 5x5) or the matrix is not a square matrix (nrow = ncol).

```{r}
library(matrixcalc)
library(Matrix)


LU_function <- function(Mat){
  if(nrow(Mat)>= 5) {
    return("Matrix too large")
  }
  
  if(is.square.matrix(Mat) == FALSE) {
    return("Matrix not square")
  }
  
  n <- dim(Mat)[1]
  L <- diag(n)
  start <- 0
  
  while (start <= dim(Mat)[1]){
  start = start+1
  i = start
  j = start
  while (i <= dim(Mat)[1]-1){
    c = Mat[i+1,j]/Mat[j,j]
    Mat[i+1,] = Mat[i+1,]-c*Mat[j,]
    L[i+1,j] = c
    i = 1+i
  }

  }
  return(L)
}


```
In order to test the function:

```{r}

testmatrix <- matrix(c(5,34,3,-32,4,66,3,1,9), nrow=3, byrow=TRUE)
(LU <- LU_function(testmatrix))

```
We can compare this with R's built in function lu.decomposition
```{r}
lu.decomposition(testmatrix)[1]

lu.decomposition(testmatrix)[[1]]==LU
```




 