---
title: "Assignment 3"
author: "Joshua Registe"
date: "9/12/2020"
output: html_document
---



# Problem Set 1

1) What is the rank of the Matrix A?

```{r}
(A<-matrix(c(1,-1,0,5,
            2,0,1,4,
            3,1,-2,-2,
            4,3,1,-2), nrow =4))


qr(A)$rank
```
2) Given an mxn matrix where m>n, what can be the maximum rank? the minimum rank? assuming that the matrix is non-zero.

<br><br>
The maximum rank of a matrix where m>n assuming non-zero matrix is n, the miminum rank will 1 as long as a matrix has at least 1 element
<br>

3) What is the rank of matrix B?

```{r}
(B<-matrix(c(1,3,2,
            2,6,4,
            1,3,2),nrow = 3))

qr(B)$rank
```

# Problem Set 2


Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work.

$$A = \begin{bmatrix} 
1  & 2 & 3\\ 
0 & 4 & 5\\ 
0 & 0 & 6\\ 
\end{bmatrix}$$

The rule of Sarros describes the characteristic equation for our given matrix as:


$$|A-\lambda I| = 0 = det \begin{bmatrix} 
1-\lambda  & 2 & 3\\ 
0 & 4-\lambda & 5\\ 
0 & 0 & 6-\lambda\\ 
\end{bmatrix}$$

meaning that our expected eigen values are 1, 4, and 6. We can compute the eigen vector as:

```{r}

A<-matrix(c(1,0,0,
            2,4,0,
            3,5,6), nrow =3)

eigen(A)


```


