---
title: "Assignment 9"
author: "Joshua Registe"
date: "10/25/2020"
output:
  html_document:
    code_folding: show
    toc: yes
    toc_float: yes
    smart: no
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, warning = F, message = F, collapse = T)
```

## Question 1 
The price of one share of stock in the Pilsdorff Beer Company (see Exercise 8.2.12) is given by $Y_n$ on the $n$th day of the year. Finn observes that the differences $X_n = Y_{n+1} - Y_n$ appear to be independent random variables with a common distribution having mean $\mu = 0$ and variance $\sigma^2 = \frac{1}{4}$. If $Y_1 = 100$, estimate the probability that $Y_{365}$ is:

(a) $\geq 100$.

(b) $\geq  110$.

(c) $\geq 120$.

-----------------------------------------------------------------------------------------

$Y_{365} = Y_1 + X_1 + X_2 + ... + X_{364}.$

$S_n = Y_{n+1} - 100$.

If $n = 364, S_{364} = Y_{365} - 100 \ \to \ Y_{365} = S_{364} + 100$.

$E[S_{364}] = n\mu = 364\cdot 0 = 0$.

Variance of $S_{364} = 364\cdot\frac{1}{4} =$ `r 364/4` $\ \to \ \sigma = \sqrt{91}$.

### (a)

$P(Y_{365} \geq 100) = P(S_{364} + 100 \geq 100) = P(S_{364} \geq 0)$.

```{r 1a}
q <- 0
mu <- 0
sd <- sqrt(91)

pnorm(q, mean = mu, sd = sd, lower.tail = F)

```

### (b)

$P(Y_{365} \geq 110) = P(S_{364} + 100 \geq 110) = P(S_{364} \geq 10)$.

where q is equal to 10 in the pnorm formula
```{r 1b}
pnorm(10, mean = mu, sd = sd, lower.tail = F)
```

### (c)

$P(Y_{365} \geq 120) = P(S_{364} + 100 \geq 120) = P(S_{364} \geq 20)$.
where q u equal to 20 in the pnorm formula
```{r 1c}

pnorm(20, mean = mu, sd = sd, lower.tail = F)
```

## Question 2
Calculate the expected value and variance of the binomial distribution using the moment generating function.

-----------------------------------------------------------------------------------------

the pmg can be described as  $\binom{n}{k}p^{n}q^{n-k}$.

And the moment generating function can be described as $M_k(t) = E[e^{tn}], \ \ t \in \mathbb{R}$.

combining the two to solve for our matrix provides 
\[M_k(t) = \sum_{k=0}^{n}e^{tn}\binom{n}{k}p^{n}q^{n-k} = \sum_{k=0}^{n}(pe^t)^{n}\binom{n}{k}q^{n-k}\]


we can simplify and take the derivative as: $M_k(t)$ wrt $t$: $M_{k}^{'}(t) = n(pe^t)(q+pe^t)^{n-1}$.

When $t = 0: E[k] = np(q+p)^{n-1} = np$.
 
using the product rule for the second moment can be described as - 
$M_k^{''}(t) = np\Big[e^{t}(pe^t + q)^{n-1}+(n-1)(pe^t + q)^{n-2}(e^tp + 0)\Big]$ 

simplifying and When $t = 0:$ provides - 

$M_k^{''}(0) = E[k^2] = np(q + p)^{n-2}(q + np) = np(q + np)$.

Using the formula $V(x) = E[x^2] - (E[x])^2:$

$V(k) = np(q + np) - n^2p^2 = npq$.

## Question 3
Calculate the expected value and variance of the exponential distribution using the moment generating function.

-----------------------------------------------------------------------------------------


Proceeding in the same manner as in question 2.

The pmf for the exponential can be descibed as $\lambda e^{-\lambda x}$.

with a moment function: \[M_x(t) = \int_{0}^{\infty}e^{tx}\lambda e^{-\lambda x} dx = \lambda \int_{0}^{\infty}e^{(t-\lambda)x} dx = \frac{\lambda}{t - \lambda}, \ \ |t| < \lambda\]

$M_{x}^{'}(t) = \frac{\lambda}{(\lambda - t)^2}$.

now When $t = 0:$ we can say - 

$E[X] = M_{x}^{'}(0) = \frac{1}{\lambda}$.

and the second moment can be described as - 

$E[X^2] = M_{x}^{''}(0) = \frac{2\lambda}{(\lambda - t)^3} = \frac{2}{\lambda^2}$.

Using the formula $V(x) = E[X^2] - (E[X])^2:$

$V(x) = \frac{2}{\lambda^2} - \frac{1}{\lambda^2} = \frac{1}{\lambda^2}$.
